### 2024-08-07 cognitive-bias.qmd
20240803 - Miller (2013)
20240807 - primacy effect bias; role of cognitive biases

As described [above](#sec-cognitiveBiasOverview), cognitive biases are systematic errors in thinking.
Cognitive biases are often due to the use of [heuristics](#sec-heuristics).
Examples of cognitive biases that result from one or more of these heuristics include:

- overconfidence bias
- confirmation bias
- in-group bias
- recency bias
- hindsight bias
- loss aversion bias
- endowment bias
- bandwagon effect bias
- Dunning–Kruger effect bias

@@

*Confirmation bias* is the tendency for people to search for, interpret, and remember information that confirms one's beliefs, as opposed to information that might disconfirm one's beliefs.
The result of confirmation bias is that people are unlikely to change their minds about something that they have a pre-existing belief about, because they tend to look only for information that supports their pre-existing beliefs.
For instance, if you believe that a particular player is a strong breakout candidate to be a sleeper, you may be more likely to pay attention to evidence that supports that the player will breakout and may be less likely to pay attention to evidence that indicates the player may struggle.

@@

### Recency Bias {#sec-cognitiveBiasesRecency}

*Recency bias* is the tendency to weigh recent events more than earlier ones.
For instance, a manager might observe that a Running Back on the waiver wire performed well in the last two games.
*Recency bias* may lead the manager to pick up the player, overvaluing their recent performance.
For instance, the manager may not have adequately weighed the player's overall season performance and the fact that the starting Running Back is returning to the lineup from injury, and that is why the player received more carries in the past two games (i.e., in place of the injured starter).

>>>>

As described [above](#sec-cognitiveBiasOverview), cognitive biases are systematic errors in thinking.
They lead us to be less accurate in our judgments and predictions.
Inaccuracy may not always be bad; several cognitive biases may serve to help people feel better about themselves, including [confirmation bias](#sec-cognitiveBiasesConfirmation), [overconfidence bias](#sec-cognitiveBiasesOverconfidence), optimism bias, self-serving bias, and actor-observer bias.
In fantasy football, however, cognitive biases and inaccurate judgments can lead to worse performance.
Cognitive biases are often due to the use of [heuristics](#sec-heuristics).

@Miller2013c describes cognitive biases in fantasy sports.
Examples of cognitive biases relevant to fantasy football that result from one or more heuristics include:

- overconfidence bias
- confirmation bias
- in-group bias
- primacy effect bias
- recency effect bias
- hindsight bias
- loss aversion bias
- endowment bias
- bandwagon effect bias
- Dunning–Kruger effect bias

@@

*Confirmation bias* is the tendency for people to search for, interpret, and remember information that confirms one's beliefs, as opposed to information that might disconfirm one's beliefs.
The result of confirmation bias is that people are unlikely to change their minds about something that they have a pre-existing belief about, because they tend to look only for information that supports their pre-existing beliefs.
For instance, if you believe that a particular player is a strong breakout candidate to be a sleeper, you may be more likely to pay attention to evidence that supports that the player will breakout and may be less likely to pay attention to evidence that indicates the player may struggle.
That is, people tend to look for confirmation that the players they want to draft (or start) are preferred by the experts [@Miller2013c].
Confirmation bias may lead to "going with your gut" (i.e., making a decision based on your intuition).

@@

### Primacy Effect Bias {#sec-cognitiveBiasesPrimacy}

*Primacy effect bias* is the tendency to recall the first information that one encounters better than information presented later on.
For instance, primacy effect bias might lead a manager to draft a player based on the first information they encountered about the player.

### Recency Effect Bias {#sec-cognitiveBiasesRecency}

*Recency effect bias* is the tendency to weigh recent events more than earlier ones.
For instance, a manager might observe that a Running Back on the waiver wire performed well in the last two games.
Recency effect bias may lead the manager to pick up the player, overvaluing their recent performance.
For instance, the manager may not have adequately weighed the player's overall season performance and the fact that the starting Running Back is returning to the lineup from injury, and that is why the player received more carries in the past two games (i.e., in place of the injured starter).

Recency effect bias seems a bit at odds with primacy effect bias; however, they can both coexist.
People tend to remember the earliest information they encounter in addition to the most recent information, and they tend to forget the information encountered in between.
